---
title: "Plotly output for motif effects"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: no
    toc_float: yes
---

```{r get_exp_name, warning = FALSE, error = FALSE, message = FALSE}
library(here)
library(plotly)
library(igraph)

#exp.name = "GAL1_7_10_query"; title = "GAL1/7/10 modifiers"
#exp.name = "GAL3_query"; title = "GAL3 introgression modifiers"
exp.name = "FLZ_resistance_modifier";title = "Modifiers of FLZ resistance"
```

```{r title, results = "asis"}
cat("##", title, "\n")
```

The following figures show the the phenotypic effects
of interacting markers compared to the additive prediction.

The points in the first two columns in each figure show the main effects
of the interacting markers. the source of the interaction is always in 
the first column, and the target is in the second. One of these is the 
query marker, and the other is the test marker depending on whether the
plot is showing the "Query as Source" interactions or the "Query as Target"
interactions.

The third column shows the additive prediction given the main effects. 
The final column shows the actual trait value of the interacting pair. 
Each interacting pair is connected by a line. The lines
are colored based on the allele of the *test marker*, since the query marker
is always the same. 

Hovering over the marker points with the mouse will reveal
where the markers are positioned in the genome, as well as genes
that are within 5 kb of the marker.

```{r near_gene_fun}
get_gene_names <- function(chr, pos, window = 5000){
  chr.locale <- which(pg_map[,1] == as.numeric(chr))
  above.min <- which(pg_map[,"start"] >= pos-window)
  below.max <- which(pg_map[,"stop"] <= pos+window)
  gene.idx <- Reduce("intersect", list(chr.locale, above.min, below.max))
  gene.name <- pg_map[gene.idx,"gene"]
  empty.fields <- which(gene.name == "")
  gene.name[empty.fields] <- pg_map[gene.idx[empty.fields],"feature"]
  return(gene.name)
}

```

```{r motif_effects}
clin.effects <- readRDS(here("Results", exp.name, "Motif_Effects.Rds"))
pg_map <- read.csv(here("Data", "pgmap.csv"))

main.types <- c("coherent", "incoherent")
int.effect <- c("aggravating", "alleviating")
int.main.sets <- rbind(cbind(rep(main.types, length(int.effect)), 
    rep(int.effect, each = length(main.types))))
type.sets <- rbind(cbind(int.main.sets, rep("", nrow(int.main.sets))),
                    cbind(int.main.sets, rep("flipped", nrow(int.main.sets))))
type.labels <- apply(type.sets, 1, function(x) paste(x, collapse = "_"))


l <- htmltools::tagList()
i = 1
for(dir in 1:length(clin.effects)){
  for(ph in 1:length(clin.effects[[dir]])){
    #ph.mat <- Reduce("rbind", lapply(1:length(clin.effects), function(x) clin.effects[[x]][[ph]]))
    ph.mat <- clin.effects[[dir]][[ph]]
    ph.mat[which(ph.mat[,"sign.flipped"] == "FALSE"),"sign.flipped"] <- ""
    ph.mat[which(ph.mat[,"sign.flipped"] == "TRUE"),"sign.flipped"] <- "flipped"
    ph.classes <- unlist(lapply(1:nrow(ph.mat), function(x) rep(paste(ph.mat[x,5:7], collapse = "_"), 4)))
    xcat <- rep(c("Main1", "Main2", "Additive", "Actual"), nrow(ph.mat))
    marker.label <- unlist(lapply(rownames(ph.mat), function(x) rep(x, 4)))
    just.marker <- sapply(strsplit(gsub("X", "", marker.label), "_"), function(x) x[1])
    marker.chr <- substr(just.marker, 1,2)
    marker.pos <- as.numeric(sapply(just.marker, function(x) substr(x, 3, nchar(x))))
    gene.list <- lapply(1:length(marker.chr), function(x) get_gene_names(marker.chr[x], marker.pos[x]))
    geneV <- sapply(gene.list, function(x) paste(x, collapse = ", "))
    marker.info <- paste0(paste0("Chr", marker.chr), "-", marker.pos, "\n", geneV)
    ypos <- as.numeric(unlist(lapply(1:nrow(ph.mat), function(x) ph.mat[x,1:4])))
    allele <- unlist(lapply(1:nrow(ph.mat), function(x) rep(ph.mat[x,8],4)))
    subj <- unlist(lapply(1:nrow(ph.mat), function(x) rep(x, 4)))

    df <- data.frame( "subject" = subj,
        "marker_chr" = marker.chr,
        "marker_pos" = marker.pos,
        "marker_info" = marker.info,
        "effect_type" = factor(xcat, levels = c("Main1", "Main2", "Additive", "Actual")),
        "effect" = ypos,
        "class" = factor(ph.classes, levels = type.labels),
        "allele" = allele)

    pl = qplot(effect_type, effect, data = df, color = I(categorical_pal(8)[as.numeric(allele)]), 
      main = paste(names(clin.effects[[dir]][ph]), names(clin.effects)[dir])) + 
      geom_line(aes(group = subject), color = categorical_pal(8)[as.numeric(allele)]) +
      aes(x = effect_type, y = effect, text = marker_info) + 
      facet_wrap(class~.)
  
    allele_colors <- categorical_pal(8)
    names(allele_colors) <- 1:8
    colScale <- scale_colour_manual(name = "allele",values = allele_colors)

    pl = qplot(effect_type, effect, data = df, color = allele,
      main = paste(names(clin.effects[[dir]][ph]), names(clin.effects)[dir])) + 
      geom_line(aes(group = subject)) +
      aes(x = effect_type, y = effect, text = marker_info) + 
      facet_wrap(class~.) +
      colScale + 
      geom_hline(aes(yintercept=0))
    #pl

    l[[i]] = as_widget(ggplotly(pl))

    i = i + 1
    
  }
} #end looping through phenotypes
```


```{r plot1, fig.width = 10, fig.height = 8}
l[[1]]
```

```{r plot2, fig.width = 10, fig.height = 10}
l[[2]]
```


```{r plot3, fig.width = 10, fig.height = 8}
l[[3]]
```

```{r plot4, fig.width = 10, fig.height = 10}
l[[4]]
```


```{r effects_across_pheno, eval = FALSE}
#do markers have the same effects across phenotypes? Different effects?
all.markers <- unique(unlist(lapply(clin.effects, 
  function(x) unlist(lapply(x, function(y) rownames(y))))))
split.markers <- strsplit(all.markers, "_")
marker.allele <- as.numeric(sapply(split.markers, function(x) x[2]))
marker.name <- sapply(split.markers, function(x) x[1])
marker.chr <- as.numeric(substr(marker.name, 2, 3))
marker.pos <- as.numeric(sapply(marker.name, function(x) substr(x, 4, nchar(x))))
marker.label <- paste0("Chr", marker.chr, "-", marker.pos)
gene.list <- lapply(1:length(marker.chr), function(x) get_gene_names(marker.chr[x], marker.pos[x]))
geneV <- sapply(gene.list, function(x) paste(x, collapse = ", "))
marker.info <- paste0(paste0("Chr", marker.chr), "-", marker.pos, "\n", geneV)

get_marker_effects <- function(marker.name, clin.effects){
  marker.effects <- lapply(clin.effects, function(x) lapply(x, function(y) y[which(rownames(y) == marker.name),]))
  effect.table <- matrix(NA, nrow = 4, ncol = 8)
  colnames(effect.table) <- colnames(clin.effects[[1]][[1]])
  row.names <- rep(NA, length(4))
  idx <- 1
  for(i in 1:length(marker.effects)){
    for(j in 1:length(marker.effects[[i]])){
      row.names[idx] <- paste(names(marker.effects)[i], names(marker.effects[[i]])[j], sep = "_")
      if(length(marker.effects[[i]][[j]] > 0)){
        effect.table[idx,] <- marker.effects[[i]][[j]]
      }
    idx = idx + 1
    }
  }
  rownames(effect.table) <- row.names
  return(effect.table)
}

all.marker.effects <- lapply(all.markers, function(x) get_marker_effects(x, clin.effects))

main.target.flz <- as.numeric(sapply(all.marker.effects, function(x) x[1,2]))
main.source.flz <- as.numeric(sapply(all.marker.effects, function(x) x[3,1]))
main.target.ypd <- as.numeric(sapply(all.marker.effects, function(x) x[2,2]))
main.source.ypd <- as.numeric(sapply(all.marker.effects, function(x) x[4,1]))

main.target.flz.add <- as.numeric(sapply(all.marker.effects, function(x) x[1,3]))
main.source.flz.add <- as.numeric(sapply(all.marker.effects, function(x) x[3,3]))
main.target.ypd.add <- as.numeric(sapply(all.marker.effects, function(x) x[2,3]))
main.source.ypd.add <- as.numeric(sapply(all.marker.effects, function(x) x[4,3]))
flz.add <- apply(cbind(main.target.flz.add, main.source.flz.add), 1, function(x) unique(x[which(!is.na(x))]))
ypd.add <- apply(cbind(main.target.ypd.add, main.source.ypd.add), 1, function(x) unique(x[which(!is.na(x))]))
#plot(cbind(flz.add, ypd.add))

main.target.flz.act <- as.numeric(sapply(all.marker.effects, function(x) x[1,4]))
main.source.flz.act <- as.numeric(sapply(all.marker.effects, function(x) x[3,4]))
main.target.ypd.act <- as.numeric(sapply(all.marker.effects, function(x) x[2,4]))
main.source.ypd.act <- as.numeric(sapply(all.marker.effects, function(x) x[4,4]))
flz.act <- apply(cbind(main.target.flz.act, main.source.flz.act), 1, function(x) unique(x[which(!is.na(x))]))
ypd.act <- apply(cbind(main.target.ypd.act, main.source.ypd.act), 1, function(x) unique(x[which(!is.na(x))]))
#plot(cbind(flz.act, ypd.act))

plot(flz.add, flz.act);abline(0,1)
plot(ypd.add, ypd.act);abline(0,1)

flz.diff <- abs(flz.act) - abs(flz.add) #positive values indicate aggravating epistasis
ypd.diff <- abs(ypd.act) - abs(ypd.add) #positive values indicate aggravating epistasis

#there is more aggravating epistasis in YPD?
boxplot(list("FLZ" = flz.diff, "YPD" = ypd.diff));abline(h = 0)
plot(flz.diff, ypd.diff);abline(0,1)

int.target.flz.type <- sapply(all.marker.effects, function(x) paste(x[1,5:7], collapse = "_"))
int.source.flz.type <- sapply(all.marker.effects, function(x) paste(x[3,5:7], collapse = "_"))
int.target.ypd.type <- sapply(all.marker.effects, function(x) paste(x[2,5:7], collapse = "_"))
int.source.ypd.type <- sapply(all.marker.effects, function(x) paste(x[4,5:7], collapse = "_"))

flz.type <- apply(cbind(int.target.flz.type, int.source.flz.type), 1, function(x) unique(x[which(x != "NA_NA_NA")]))
ypd.type <- apply(cbind(int.target.ypd.type, int.source.ypd.type), 1, function(x) unique(x[which(x != "NA_NA_NA")]))

any.main.flz <- apply(cbind(main.target.flz, main.source.flz), 1, function(x) unique(x[which(!is.na(x))]))
any.main.ypd <- apply(cbind(main.target.ypd, main.source.ypd), 1, function(x) unique(x[which(!is.na(x))]))

all.main <- cbind(any.main.flz, any.main.ypd)
colnames(all.main) <- c("FLZ", "YPD")
rownames(all.main) <- all.markers

main.df <- data.frame("marker_label" = marker.info, "FLZ" = all.main[,1], 
  "YPD" = all.main[,2], "allele" = as.factor(marker.allele), 
  "FLZ_add" = flz.add, "FLZ_act" = flz.act, "FLZ_diff" = flz.diff, 
  "YPD_add" = ypd.add, "YPD_act" = ypd.act, "YPD_diff" = ypd.diff, 
  "FLZ_type" = flz.type, "YPD_type" = ypd.type)

main.pl <- qplot(FLZ, YPD, data = main.df, color = allele, main = "Main Effects in Each Trait")+ 
  colScale + 
  geom_hline(aes(yintercept=0)) +
  geom_vline(aes(xintercept=0)) +
  aes(x = FLZ, y = YPD, text = marker_label) +
  annotate("text", x = 0.52, y = 0.35, label = "Improved\nGrowth\nOverall", adj = 1) +
  annotate("text", x = -0.35, y = -0.4, label = "Reduced Growth\nOverall", adj = 0) +
  annotate("text", x = 0.52, y = -0.4, label = "Drug resistance at\nexpense of overall growth", adj = 1) + 
  annotate("text", x = -0.35, y = 0.38, label = "Drug susceptibility", adj = 0)

main.pl  
ggplotly(main.pl)

act.pl <- qplot(FLZ_act, YPD_act, data = main.df, color = allele, main = "Actual Growth for Interaction") +
  colScale + 
  geom_hline(aes(yintercept=0)) +
  geom_vline(aes(xintercept=0)) +
  aes(x = FLZ_act, y = YPD_act, text = marker_label)
act.pl
ggplotly(act.pl)


int.pl <- qplot(FLZ_diff, YPD_diff, data = main.df, color = allele, main = "Degree of Aggravation/Alleviation") + 
  colScale + 
  geom_hline(aes(yintercept=0)) +
  geom_vline(aes(xintercept=0)) +
  aes(x = FLZ_diff, y = YPD_diff, text = marker_label)
  
ggplotly(int.pl)


flz.idx <- grep("FLZ", colnames(main.df))
ypd.idx <- grep("YPD", colnames(main.df))

main.df[1,flz.idx]
main.df[1,ypd.idx]

ylim <- c(min((main.df[,c(flz.idx[1:4], ypd.idx[1:4])])), max((main.df[,c(flz.idx[1:4], ypd.idx[1:4])])))

pdf("~/Desktop/test.pdf", width = 9, height = 9)
par(mfrow = c(3,3))
for(i in 1:nrow(main.df)){
  plot.new()
  plot.window(xlim = c(1,3), ylim = ylim)
  barplot(rbind(unlist(main.df[i,flz.idx[1:3]]), unlist(main.df[i,ypd.idx[1:3]])), beside = TRUE)
  points(1:3, , type = "b")
  points(1:3, , type = "b", col = "red")
  abline(h = 0)
  axis(1);axis(2)
}
dev.off()
```