---
title: "Query CAPE Output"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: no
    toc_float: yes
---

```{r get_exp_name, warning = FALSE, error = FALSE, message = FALSE}
library(here)
library(plotly)

exp.name = "GAL1_7_10_query"
exp.name = "GAL3_query"
```

The following figures show the deviation from additivity when the 
query locus, `r exp.name`, interacts with each marker in the genome.
Only interactions that were significant in CAPE are shown. 

Interactions for each chromosome are shown in a separate plot.
Each plot is split into a top portion showing the interactions
in which the query locus was the source of the interaction, and a
bottom portion in which the query locus was the target of the 
interaction.

The dot of each lollipop is colored by the allele participating 
in the interaction. The stem of each lollipop is colored by the
trait that the interaction affected. The length of the lollipop
indicates the amount that the trait deviated from the additive
expectation when the query allele and the test allele were 
found in the same individuals.

By hovering with the mouse above the lollipops, you can see
the names of genes that are within 5kb of each marker, along
with the precise location of the marker.

I haven't yet figured out how to fix the $y$ axes for these plots
yet, so they are all on individual $y$ axes.

```{r variant_effects, fig.width = 10, fig.height = 8}
var_effects <- readRDS(here("Results", exp.name, "Plotly_variant_effects.RDS"))

l <- htmltools::tagList()
for (i in 1:length(var_effects)){ 
  l[[i]] = as_widget(var_effects[[i]])
}

l
```