---
title: "Query CAPE Analysis"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---

The purpose of this workflow is to analyze results from Query_CAPE.Rmd
```{r get_exp_name}
library(here)
exp.name <- "GAL_query"
results.dir <- here("Results", exp.name)
is.interactive = FALSE
#is.interactive = TRUE
```

The results here pertain to the directory `r exp.name`.

```{r load_code}
all.fun <- list.files(here("Code"), full.names = TRUE, pattern = ".R")
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r load_cape, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
#load local version of cape code
needed.libraries <- c("pheatmap")
load_libraries(needed.libraries, personal.library = TRUE)
load_latest_cape(here("Code", "cape"), personal.library = TRUE)
```

## Load data

```{r read_data}
final_obj <- readRDS(file.path(results.dir, "cross.RDS"))
geno_obj <- readRDS(file.path(results.dir, "cross_geno.RDS"))
```

The following barplot shows the total number of interactions per
chromosome.

```{r var_inf}
var_inf <- write_variant_influences(final_obj, p_or_q = 1e-6, 
  include_main_effects = FALSE, write_file = FALSE)
all_chr <- c(var_inf[,2], var_inf[,5])

barplot(table(all_chr[which(all_chr != 0)]), 
  main = "Number of Interactions per Chromosome")
```

The following plot shows the interaction counts along the 
genome in a sliding window.

```{r interactions_along_genome}
u_chr <- 1:16

interaction_count <- function(u_chr, marker_names, window.size = 1e4, gap.size = 5e3){
  split_marker <- strsplit(marker_names, "_")
  marker_chr <- sapply(split_marker, function(x) as.numeric(substr(x[[1]], 2,3)))
  marker_pos <- sapply(split_marker, function(x) as.numeric(substr(x[[1]], 4, nchar(x))))
  chr_idx <- lapply(u_chr, function(x) which(marker_chr == x))
  chr_pos <- lapply(chr_idx, function(x) sort(marker_pos[x]))
  windows <- lapply(chr_pos, function(x) if(length(x) > 0){sliding.window.el(elV = 1:ceiling(max(x)), 
    window.size = window.size, gap.size = gap.size)}else{NA})
  counts <- lapply(1:length(windows), function(x) sapply(1:length(windows[[x]]), 
    function(y) if(length(windows[[x]][[y]]) > 0){length(intersect(which(chr_pos[[x]] >= min(windows[[x]][[y]])), 
    which(chr_pos[[x]] <= max(windows[[x]][[y]]))))}else{NA}))
  count_pos <- lapply(windows, function(x) sapply(x, mean))
  result <- list("positions" = count_pos, "counts" = counts)
  return(result)
}

plot_counts <- function(count_result, plot.label = ""){
  int_counts <- count_result$counts
  int_pos <- count_result$positions
  chr <- 1:length(int_counts)
  
  global.max <- max(unlist(int_counts), na.rm = TRUE)
  layout.mat <- matrix(chr, nrow = 1)
  layout(layout.mat)

  for(ch in 1:length(chr)){
    par(mar = c(4,0,4,0))
    if(ch == 1){
      par(mar = c(4,4,4,0))
    }
    if(ch == max(u_chr)){
      par(mar = c(4,0,4,2))
    }
    if(!all(is.na(int_pos[[ch]]))){
      plot(int_pos[[ch]], int_counts[[ch]], axes = FALSE, 
        type = "h", xlab = "", ylab = "", ylim = c(0, global.max))
    }else{
      plot.new()
      plot.window(xlim = c(0,1), ylim = c(0, global.max))
      abline(h = 0)
    }
    mtext(paste0("Chr", ch), side = 1, line = 0)
    if(ch == 1){axis(2)}
  }
  mtext("Interaction Counts", side = 2, line = -2, outer = TRUE)
  mtext(plot.label, side = 3, line = -2, outer = TRUE)
}

plot_two_counts <- function(count_results1, count_results2, plot.label = ""){

  int_counts1 <- count_results1$counts
  int_pos1 <- count_results1$positions
  int_counts2 <- count_results2$counts
  int_pos2 <- count_results2$positions
  chr <- 1:max(c(length(int_counts1), length(int_counts2)))
  
  global.max <- max(c(unlist(int_counts1), unlist(int_counts2)), na.rm = TRUE)
  layout.mat <- matrix(chr, nrow = 1)
  layout(layout.mat)

  for(ch in 1:length(chr)){
    par(mar = c(4,0,4,0))
    if(ch == 1){
      par(mar = c(4,4,4,0))
    }
    if(ch == max(u_chr)){
      par(mar = c(4,0,4,2))
    }
    plot.new()

    if(!all(is.na(int_pos1[[ch]])) && !all(is.na(int_pos2[[ch]]))){
      plot.window(xlim = c(0, max(c(int_pos1[[ch]], int_pos2[[ch]]))), 
        ylim = c(global.max*-1, global.max))
      if(!all(is.na(int_pos1[[ch]]))){
        points(int_pos1[[ch]], int_counts1[[ch]], type = "h", col = "#bf5b17")
      }
      if(!all(is.na(int_pos2[[ch]]))){
      points(int_pos2[[ch]], int_counts2[[ch]]*-1, type = "h", col = "#386cb0")    
      }
    }else{
        plot.window(xlim = c(0,1), ylim = c(-global.max, global.max))
        abline(h = 0)
    }
    mtext(paste0("Chr", ch), side = 1, line = 0)
    if(ch == 1){axis(2)}
  }
  mtext("Interaction Counts", side = 2, line = -2, outer = TRUE)
  mtext(plot.label, side = 3, line = -2, outer = TRUE)

}
```

The following plots show how the source markers are distributed
across the genome. These are the markers that target the GAL
introgression.

```{r source_dist, fig.width = 10, fig.height = 5}
source_markers <- var_inf[,"Source"]
source_markers <- source_markers[which(source_markers != "query")]
source_counts <- interaction_count(u_chr, source_markers, 1e4, 5e3)
if(is.interactive){quartz(width = 10, height = 5)}
plot_counts(source_counts, plot.label = "Source Marker Positions")
```

The following plots show how the target markers are distributed
across the genome. These are the markers that are targeted by 
the GAL introgression.


```{r target_dist, fig.width = 10, fig.height = 5}
target_markers <- var_inf[,"Target"]
target_markers <- target_markers[which(target_markers != "query")]
target_counts <- interaction_count(u_chr, target_markers, 1e4, 5e3)
if(is.interactive){quartz(width = 10, height = 5)}
plot_counts(target_counts, plot.label = "Target Marker Positions")
```

The following plot shows a direct comparison of the distributions of the
source markers (top, brown) and the target markers (bottom blue).

```{r comp, fig.width = 10, fig.height = 5}
plot_two_counts(count_results1 = source_counts, count_results2 = target_counts)
```