---
title: "Query CAPE Analysis"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---

The purpose of this workflow is to analyze results from
multiple runs of Query_CAPE.Rmd

```{r get_exp_name}
library(here)
exp.names <- c("GAL3_query", "GAL4_query")
is.interactive = FALSE
#is.interactive = TRUE
u_chr <- 1:16
```

The results here pertain to the directory `r exp.names`.

```{r load_code}
all.fun <- list.files(here("Code"), full.names = TRUE, pattern = ".R")
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r load_cape, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
#load local version of cape code
needed.libraries <- c("pheatmap")
load_libraries(needed.libraries, personal.library = TRUE)
load_latest_cape(here("Code", "cape"), personal.library = TRUE)
```

## Load data

The following barplots show the total number of interactions per
chromosome.

```{r var_inf}
var_inf <- lapply(exp.names, 
  function(x) as.matrix(read.csv(here("Results", x, "Variant_Influences_Interactions.csv"))))
```

```{r int_per_chrom}
int_per_chrom <- function(var.inf.table, chrom = 1:16, p.or.q = 1e-6, 
    plot.results = FALSE, plot.label = ""){
  sig.idx <- which(as.numeric(var.inf.table[,"p_adjusted"]) <= p.or.q)
  source.query.idx <- which(var.inf.table[,"Source"] == "query")
  target.query.idx <- which(var.inf.table[,"Target"] == "query")

  source.query.table <- var.inf.table[intersect(sig.idx,source.query.idx),]
  target.query.table <- var.inf.table[intersect(sig.idx,target.query.idx),]

  source.chrom.idx <- lapply(chrom, function(x) which(as.numeric(source.query.table[,5]) == x))
  target.chrom.idx <- lapply(chrom, function(x) which(as.numeric(target.query.table[,2]) == x))

  source.chrom.count <- sapply(source.chrom.idx, length)
  target.chrom.count <- sapply(target.chrom.idx, length)
  counts.by.chrom <- rbind(source.chrom.count, target.chrom.count)
  colnames(counts.by.chrom) <- chrom
  rownames(counts.by.chrom) <- c("Query as Source", "Query as Target")

  if(plot.results){
    if(is.interactive){quartz(width = 10, height = 5)}
    barplot(counts.by.chrom, beside = TRUE, 
    col = c("#7fc97f", "#beaed4"), main = plot.label)
    legend("topright", fill = c("#7fc97f", "#beaed4"), 
      legend = c("Query as Source", "Query as Target"))
  }

  source.chrom.effect <- lapply(source.chrom.idx, function(x) as.numeric(source.query.table[x,"Effect"]))
  names(source.chrom.effect) <- chrom
  target.chrom.effect <- lapply(target.chrom.idx, function(x) as.numeric(target.query.table[x,"Effect"]))
  names(target.chrom.effect) <- chrom
  all.effects <- list(source.chrom.effect, target.chrom.effect)
  names(all.effects) <- c("Query as Source", "Query as Target")
  
  if(plot.results){
    if(is.interactive){quartz(width = 10, height = 5)}
    plot.grouped.boxes(all.effects, print.vals = NA, main = plot.label)
    abline(h = 0)
  }

  all.results <- list("Counts_by_Chromosome" = counts.by.chrom, 
  "Effects_by_Chrom" = all.effects)
  invisible(all.results)

}

for(i in 1:length(var_inf)){
  int_per_chrom(var_inf[[i]], plot.results = TRUE, plot.label = exp.names[i])
}
```

The following plot shows the interaction counts along the 
genome in a sliding window.

```{r interactions_along_genome}

interaction_count <- function(u_chr, marker_names, window.size = 1e4, gap.size = 5e3){
  split_marker <- strsplit(marker_names, "_")
  marker_chr <- sapply(split_marker, function(x) as.numeric(substr(x[[1]], 2,3)))
  marker_pos <- sapply(split_marker, function(x) as.numeric(substr(x[[1]], 4, nchar(x))))
  chr_idx <- lapply(u_chr, function(x) which(marker_chr == x))
  chr_pos <- lapply(chr_idx, function(x) sort(marker_pos[x]))
  windows <- lapply(chr_pos, function(x) if(length(x) > 0){sliding.window.el(elV = 1:ceiling(max(x)), 
    window.size = window.size, gap.size = gap.size)}else{NA})
  counts <- lapply(1:length(windows), function(x) sapply(1:length(windows[[x]]), 
    function(y) if(length(windows[[x]][[y]]) > 0){length(intersect(which(chr_pos[[x]] >= min(windows[[x]][[y]])), 
    which(chr_pos[[x]] <= max(windows[[x]][[y]]))))}else{NA}))
  count_pos <- lapply(windows, function(x) sapply(x, mean))
  result <- list("positions" = count_pos, "counts" = counts)
  return(result)
}

plot_counts <- function(count_result, plot.label = ""){
  int_counts <- count_result$counts
  int_pos <- count_result$positions
  chr <- 1:length(int_counts)
  
  global.max <- max(unlist(int_counts), na.rm = TRUE)
  layout.mat <- matrix(chr, nrow = 1)
  layout(layout.mat)

  for(ch in 1:length(chr)){
    par(mar = c(4,0,4,0))
    if(ch == 1){
      par(mar = c(4,4,4,0))
    }
    if(ch == max(u_chr)){
      par(mar = c(4,0,4,2))
    }
    if(!all(is.na(int_pos[[ch]]))){
      plot(int_pos[[ch]], int_counts[[ch]], axes = FALSE, 
        type = "h", xlab = "", ylab = "", ylim = c(0, global.max))
    }else{
      plot.new()
      plot.window(xlim = c(0,1), ylim = c(0, global.max))
      abline(h = 0)
    }
    mtext(paste0("Chr", ch), side = 1, line = 0)
    if(ch == 1){axis(2)}
  }
  mtext("Interaction Counts", side = 2, line = -2, outer = TRUE)
  mtext(plot.label, side = 3, line = -2, outer = TRUE)
}

plot_two_counts <- function(count_results1, count_results2, plot.label = ""){

  int_counts1 <- count_results1$counts
  int_pos1 <- count_results1$positions
  int_counts2 <- count_results2$counts
  int_pos2 <- count_results2$positions
  chr <- 1:max(c(length(int_counts1), length(int_counts2)))
  
  global.max <- max(c(unlist(int_counts1), unlist(int_counts2)), na.rm = TRUE)
  layout.mat <- matrix(chr, nrow = 1)
  layout(layout.mat)

  for(ch in 1:length(chr)){
    par(mar = c(4,0,4,0))
    if(ch == 1){
      par(mar = c(4,4,4,0))
    }
    if(ch == max(u_chr)){
      par(mar = c(4,0,4,2))
    }
    plot.new()

    if(!all(is.na(int_pos1[[ch]])) && !all(is.na(int_pos2[[ch]]))){
      plot.window(xlim = c(0, max(c(int_pos1[[ch]], int_pos2[[ch]]))), 
        ylim = c(global.max*-1, global.max))
      if(!all(is.na(int_pos1[[ch]]))){
        points(int_pos1[[ch]], int_counts1[[ch]], type = "h", col = "#bf5b17")
      }
      if(!all(is.na(int_pos2[[ch]]))){
      points(int_pos2[[ch]], int_counts2[[ch]]*-1, type = "h", col = "#386cb0")    
      }
    }else{
        plot.window(xlim = c(0,1), ylim = c(-global.max, global.max))
        abline(h = 0)
    }
    mtext(paste0("Chr", ch), side = 1, line = 0)
    if(ch == 1){axis(2)}
  }
  mtext("Interaction Counts", side = 2, line = -2, outer = TRUE)
  mtext(plot.label, side = 3, line = -2, outer = TRUE)

}
```

The following plots show how the source markers are distributed
across the genome. 

```{r source_dist, fig.width = 10, fig.height = 5}
window.size <- 1e4; gap.size = window.size/2
all.source.counts <- vector(mode = "list", length = length(var_inf))
names(all.source.counts) <- exp.names
for(i in 1:length(var_inf)){
  source_markers <- var_inf[[i]][,"Source"]
  source_markers <- source_markers[which(source_markers != "query")]
  source_counts <- interaction_count(u_chr, marker_names = source_markers, 
    window.size = window.size, gap.size = gap.size)
  if(is.interactive){quartz(width = 10, height = 5)}
  plot_counts(source_counts, plot.label = paste(exp.names[i], "Source Marker Positions"))
  all.source.counts[[i]] <- source_counts
}
```

The following plots show how the target markers are distributed
across the genome. These are the markers that are targeted by 
the GAL introgression.


```{r target_dist, fig.width = 10, fig.height = 5}
all.target.counts <- vector(mode = "list", length = length(exp.names))
names(all.target.counts) <- exp.names
for(i in 1:length(var_inf)){
  target_markers <- var_inf[[i]][,"Target"]
  target_markers <- target_markers[which(target_markers != "query")]
  target_counts <- interaction_count(u_chr, target_markers, window.size, gap.size)
  if(is.interactive){quartz(width = 10, height = 5)}
  plot_counts(target_counts, plot.label = "Target Marker Positions")
  all.target.counts[[i]] <- target_counts
}
```

The following plot shows a direct comparison of the distributions of the
source markers (top, brown) and the target markers (bottom blue).

```{r comp, fig.width = 10, fig.height = 5}
for(i in 1:length(exp.names)){
  if(is.interactive){quartz(width = 10, height = 5)}
  plot_two_counts(count_results1 = all.source.counts[[i]], 
  count_results2 = all.target.counts[[i]], plot.label = exp.names[i])
}
```

## Effect Sizes

```{r effect_sizes}
for(i in 1:length(var_inf)){
  query.as.source <- which(var_inf[[i]][,"Source"] == "query")
  query.as.target <- which(var_inf[[i]][,"Target"] == "query")
  source.effects <- as.numeric(var_inf[[i]][query.as.source,"Effect"])
  target.effects <- as.numeric(var_inf[[i]][query.as.target,"Effect"])

  if(is.interactive){quartz()}
  boxplot(list(source.effects,target.effects), 
    names = c("Query as Source", "Query as Target"), main = exp.names[i])
  abline(h = 0)
}
```

## Network Effects

We would eventually like to build a network of effects
across multiple query cape runs. How many of the interactions
of one query are present in the other?

```{r overlapping_interactions, fig.width = 8, fig.height = 4}
all.interactors <- lapply(var_inf, 
  function(x) list(unique(x[which(x[,"Source"] != "query"),"Source"]), 
  unique(x[which(x[,"Target"] != "query"),"Target"])))
names(all.interactors) <- exp.names

query_pairs <- pair.matrix(1:length(exp.names))

for(i in 1:nrow(query_pairs)){
  exp1 <- query_pairs[i,1]
  exp2 <- query_pairs[i,2]
  common.query.sources <- intersect(all.interactors[[exp1]][[1]], all.interactors[[exp2]][[1]])
  common.source.effects <- cbind(as.numeric(var_inf[[exp1]][match(common.query.sources, var_inf[[exp1]][,"Source"]),"Effect"]),
    as.numeric(var_inf[[exp2]][match(common.query.sources, var_inf[[exp2]][,"Source"]),"Effect"]))

  common.query.targets <- intersect(all.interactors[[exp1]][[2]], all.interactors[[exp2]][[2]])
  common.target.effects <- cbind(as.numeric(var_inf[[exp1]][match(common.query.targets, var_inf[[exp1]][,"Target"]),"Effect"]),
    as.numeric(var_inf[[exp2]][match(common.query.targets, var_inf[[exp2]][,"Target"]),"Effect"]))

  if(is.interactive){quartz(width = 8, height = 4)}
  par(mfrow = c(1,2))
  plot.min <- min(c(min(common.source.effects), min(common.target.effects)))
  plot.max <- max(c(max(common.source.effects), max(common.target.effects)))
  
  plot(common.target.effects, xlab = exp.names[exp1], ylab = exp.names[exp2],
  xlim = c(plot.min, plot.max), ylim = c(plot.min, plot.max), main = "Query as Source")
  abline(h = 0, v = 0)

  plot(common.source.effects, xlab = exp.names[exp1], ylab = exp.names[exp2],
  xlim = c(plot.min, plot.max), ylim = c(plot.min, plot.max), main = "Query as Target")
  abline(h = 0, v = 0)

  mtext(paste(exp.names[exp1], "vs." exp.names[exp2]), side = 3, line = -1.5, outer = TRUE)
}

```