---
title: "CAPE results"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: show
    collapsed: no
    toc: yes
    toc_float: yes
params:
    results_dir: here("tests", "testthat", "results")
---

Set the working directory below and run this workflow to see figures of results
with explanations.

```{r set_exp}
library("here")
exp.name <- "chr15"
#exp.name <- "pathway_two_genes"
results_dir <- here("Results", exp.name)
```


```{r load_code}
all.fun <- list.files(here("Code"), full.names = TRUE, pattern = ".R")
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r load_cape, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
#load local version of cape code
needed.libraries <- c("pheatmap", "knitr", "DT")
load_libraries(needed.libraries, personal.library = TRUE)
load_latest_cape(here("Code", "cape"), personal.library = TRUE)
```

```{r load_code_redo}
all.fun <- list.files(here("Code"), full.names = TRUE, pattern = ".R")
for(i in 1:length(all.fun)){source(all.fun[i])}
```


## Introduction

This workflow summarizes results from a specified cape run. 

## Trait Description

The following plots show the correlations between traits used in this CAPE analysis.

```{r read_effects}
data_obj <- readRDS(file.path(results_dir, "cross.RDS"))
geno_obj <- readRDS(file.path(results_dir, "cross_geno.RDS"))
var.inf <- as.matrix(read.csv(file.path(results_dir, "Variant_Influences_Interactions.csv"), stringsAsFactors = FALSE))
```

```{r pheno_cor}
plot_pheno_cor(data_obj)
```


## Trait Decomposition

The first step in CAPE is often to perform a singular value decomposition (SVD) 
on the trait matrix. This step decomposes the traits into decorrelated, composite 
traits called eigentraits. The reasoning behind this step is that correlated 
components of related traits will be concentrated into individual eigentraits
thereby making these components easier to map. 

We plot this decomposition to show how each trait contributes to each eigentrait,
as well as the variance explained by each eigentrait. 

In the following plot, traits are shown along the $y$-axis and eigentraits
are shown along the $x$-axis. Each cell of the heat map indicates how much 
each trait contributes to each eigentrait. Positive contributions are shown 
in brown and negative contributions are shown in blue. Darker colors indicate 
stronger contributions. This heat map allows us to see which traits are compared 
or contrasted in the individual eigentraits. 

The gray bars along the top of the heat map indicate how much overall trait 
variance is explained by each eigentrait. They are always listed in order of 
the amount of variance explained. 

```{r svd}
plot_svd(data_obj)
```

## Single Locus Effects {.tabset .tabset-fade .tabset-pills}

Before performing the pairwise marker scans, we assess the association between
each individual marker and each trait or eigentrait using marker regression. 
These results are shown in the plots below.

The singlescan plots show how strongly each marker is associated with each 
trait. The strength of these main effects can be uses as a marker-selection
criterion for performing the pair-wise marker scans. When 
*marker_selection_method* is set to *top_effects*, the markers with the 
largest main effects are selected for the pair-wise marker scan regardless of 
significance. 

The single-locus effects are displayed in two different types of plots. 
One plot shows the overall locus effect, as well as the individual allele 
effects. For a two-parent cross, these are identical. However, for a multi-parent
cross, such as the Diversity Outbred (DO) or Collaborative Cross (CC) mice,
the allele effects will show how each parental haplotype affects the trait. 

In CAPE we use one parental haplotype as the reference. Thus all haplotype 
effects shown are relative to that reference. The reference in a DO/CC population
is typically the B6 (B) allele. 

The effects in the plots below are shown as lines, but this, and many other 
plotting parameters can be changed in *plot_singlescan()*.


```{r singlescan_effects, results = "asis", echo = FALSE}
fig.files <- list.files(path = results_dir, pattern = "Effects.jpg")
fig.names <- gsub(".jpg", "", fig.files)
for(i in 1:length(fig.files)){
    cat("###", fig.names[i], "\n")
    fig.line <- paste0("![", fig.names[i], "](", file.path(results_dir, fig.files[i]), ")\n\n")
    cat(fig.line)
}
```


## Standardized Single Locus Effects {.tabset .tabset-fade .tabset-pills}


The other way we display the single-locus effects is as standardized effects, 
which shows how significant each effect is, rather than the size and the 
direction of the effects. These standardized effects are shown below.

```{r singlescan_std, results = "asis", echo = FALSE}
fig.files <- list.files(path = results_dir, pattern = "Standardized.jpg")
fig.names <- gsub(".jpg", "", fig.files)
for(i in 1:length(fig.files)){
    cat("###", fig.names[i], "\n")
    fig.line <- paste0("![", fig.names[i], "](", file.path(results_dir, fig.files[i]), ")\n\n")
    cat(fig.line)
}
```


## Variant Influences 

The Combined Analysis of Pleiotropy and Epistasis results in a network
of directed genetic interactions. These results are plotted in multiple
different ways. 

### Variant influences as a matrix

One way to show the directed influences is in a matrix. In the following
plot, genetic markers are listed both on the $x$-axis and the $y$-axis.
The $x$-axis shows the markers as sources. Genetic interactions are 
directed from these source markers to the target markers on the $y$-axis. 
They are shown as either positive (brown) and negative (blue). Main effects
are shown along the right-hand side of the plot in extra columns.
These main effects line up with the source markers. The main effects
are also shown as positive (brown), or negative (blue). Main effects
are shown in terms of the original traits if *transform_to_phenospace*
is set to TRUE, or the eigentraits if *transform_to_phenospace* is set
to FALSE.

By default, cape does not test marker pairs that are correlated with a
Pearson $R >= 0.5$. Entries for pairs that are not tested are shown in 
gray. Gray blocks indicate areas of highly correlated markers. This 
matrix type of plot is good for seeing overall patterns of interactions
for dense matrices, or small numbers of markers. It is not very good for
visualizing sparse matrices derived from large numbers of markers.

```{r variant_influences, fig.width = 7, fig.height = 7}
plot_variant_influences(data_obj)
```


### Circular network plot

Genetic interactions can also be visualized in a circular plot. In these plots,
the chromosomes are plotted in a circle. The sizes are not representative of 
the true chromosome sizes. Main effects are shown in concentric circles around
the chromosomes, one trait per circle. Main effects are depicted as colored 
points or bars in this plot. For example, a brown bar at chromosome 7 for 
trait 1 means that there is a locus on chromosome 7 with a positive main 
effect on trait 1.

Genetic interactions in this plot are shown as colored arrows between 
genetic locations. Arrows show the direction of the influence, and the 
color indicates whether the source variant enhances (brown) or suppresses
(blue) the effects of the target marker. 

Main effects and interaction effects are also shown for covariates, which
are plotted in the circle before chromosome 1. 

This type of plot is good for seeing patterns in the physical location of 
genetic interactions in sparse networks. It is not good for visualizing 
networks with many interactions.

```{r circular_network, results = "asis", echo = FALSE}
fig.file <- file.path(results_dir, "Network_Circular.jpg")
fig.line <- paste0("![Circular Network](", fig.file, ")\n")
cat(fig.line)
```


### Network view

Finally, the cape interaction network is shown in a more classic network view. 
In this view, each genetic marker is depicted as a circle. Its main effects
are shown in slices of the pie as either positive (brown) or negative (blue).

Interactions are shown as arrows between linkage blocks. We cluster markers
into linkage blocks based on pairwise correlation. Network nodes are labeled
by chromosome followed by a number indicating the linkage block. Linkage blocks
can contain one or more markers. Covariates are also included in these network 
views, and have bold outlines. 

This type of plot is good for seeing patterns in network connectivity, such as 
clusters of enhancing or suppressing interactions, or nodes that serve
as network hubs. Genomic position information is lost in this view and thus 
this type of network is not good for visualizing the physical location of 
interacting markers. 

```{r network_layout, results = "asis", echo = FALSE}
fig.file <- file.path(results_dir, "Network_View.jpg")
fig.line <- paste0("![Network Layour](", fig.file, ")\n")
cat(fig.line)
```

## Variant influences tables {.tabset .tabset-fade .tabset-pills}

Another useful output of cape is the variant influences table. This table
contains all the specific information about interactions up to an adjusted
$p$ value of 0.2. The run_cape() pipeline writes two tables: 
Variant_Influences.csv and Variant_Influences_Interactions.csv.

The first table, Variant_Influences.csv, contains both genetic interactions
and main effects. Interactions are listed by their source marker and target 
marker. If the effect listed is a main effect, the target listed is a trait
rather than a marker. Because all effects are assessed from pairwise models,
even the main effects are listed with a conditioning marker. The main effect
of a marker is the largest effect that marker had conditioned on another 
marker. The conditioning marker and its positional information are listed 
in the case of main effects. Covariates can also have interactions and 
main effects. In this case, they are listed as the target our source marker.

These tables can be used to identify and plot specific interactions.
They are sorted by standardized effect size. Interactions and main effects 
can be plotted using the function plot_effects().

The following tables show the variant influences both with and without
main effects. 

### All Influences

```{r var_inf_table, echo = FALSE}
var.inf.table <- as.matrix(read.csv(file.path(results_dir, "Variant_Influences.csv"), stringsAsFactors = FALSE))
var.inf.table[,1] <- gsub("X", "Chr", var.inf.table[,1])
var.inf.table[,4] <- gsub("X", "Chr", var.inf.table[,4])
datatable(var.inf.table)
```

### Just Interactions

```{r var_int_table, echo = FALSE}
var.int.table <- as.matrix(read.csv(file.path(results_dir, "Variant_Influences_Interactions.csv"), stringsAsFactors = FALSE))
var.int.table[,1] <- gsub("X", "Chr", var.int.table[,1])
var.int.table[,4] <- gsub("X", "Chr", var.int.table[,4])
datatable(var.int.table)
```

## Motif Effects {.tabset .tabset-fade .tabset-pills}

The following plots show the phenotypic effects of individual
motifs compared to the additive expectation.

```{r parse_inf, fig.width = 8, fig.height = 8}
pheno.effect.mats <- pheno.effects.DO(data.obj = data_obj, 
  geno.obj = geno_obj, color.scheme = "other", geno.coding = "Dominant")
```

The following tables show detailed information about the interactions above.

```{r motif_tables, results = "asis"}
for(i in 1:length(pheno.effect.mats)){
  if(length(pheno.effect.mats[[i]]) > 0){
    cat("###", names(pheno.effect.mats)[i], "\n")
    df <- cbind(pheno.effect.mats[[i]][,c(1:3,8),drop=FALSE], 
      matrix(apply(pheno.effect.mats[[i]][,4:7,drop=FALSE], 2, 
      function(x) round(as.numeric(x), 2)), ncol = 4))
    colnames(df) <- c(colnames(pheno.effect.mats[[i]])[c(1:3,8)], colnames(pheno.effect.mats[[i]])[c(4:7)])
    rownames(df) <- NULL
    print(kable(df))
    cat("\n\n")
  }
}

#library(DT)
#datatable(df)
```
