---
title: "Yeast Run CAPE"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---

The purpose of this workflow is to run cape on the yeast data.
The experiment name specifies which parameter file to use.

```{r get_exp_name}
library(here)

args <- commandArgs(trailingOnly=T)
exp.name = args[1]
delete_previous <- as.logical(args[2])

if(is.na(exp.name)){
    exp.name <- "chr15_gal_covar"
    delete_previous <- FALSE
}

results.dir <- here("Results", exp.name)
remove_chr9 = TRUE
```

```{r delete_old_data}
if(delete_previous){
    patterns <- c(".txt", ".csv", ".RDS", ".jpg", ".pdf")
    to.delete <- unlist(lapply(patterns, function(x) list.files(results.dir, pattern = x, full.names = TRUE)))
    unlink(to.delete)
}
```

```{r load_code}
all.fun <- list.files(here("Code"), full.names = TRUE, pattern = ".R")
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r load_cape, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
#load local version of cape code
needed.libraries <- c("pheatmap")
load_libraries(needed.libraries, personal.library = TRUE)
load_latest_cape(here("Code", "cape"), personal.library = TRUE)
```

Load file and run cape.

```{r read_data}
pop <- read_population(filename = here("Data", "scaffold_72h_data_rqtl1_form_new_11_16_21.csv"),
  id_col = 1, na_strings = NA)
#str(pop)

cross_obj <- cape2mpp(pop)
pheno_obj <- cross_obj[[1]]

#log transform the phenotypes
pheno_obj$pheno <- apply(pheno_obj$pheno, 2, log10)
#mean center
pheno_obj$pheno <- apply(pheno_obj$pheno, 2, function(x) x - mean(x, na.rm = TRUE))

if(remove_chr9){
  geno <- cross_obj[[2]]$geno[,,which(cross_obj[[1]]$chromosome != 9)]
}else{
  geno <- cross_obj[[2]]$geno
}

#str(pheno_obj)
geno_obj <- geno
#str(geno_obj)

```


```{r run_cape}
final.file <- file.path(results.dir, "cross.RDS")

if(!file.exists(final.file)){

  final_obj <- run_cape(pheno_obj, geno_obj, 
    results_file = "cross.RDS", p_or_q = 0.05, n_cores = 4, 
    initialize_only = FALSE, verbose = TRUE, run_parallel = FALSE, 
    param_file = file.path(results.dir, "cape.parameters.yml"), 
    yaml_params = NULL, results_path = results.dir)
}else{
  final_obj <- readRDS(final.file)
  geno_obj <- readRDS(file.path(results.dir, "cross_geno.RDS"))
  pairscan_obj <- readRDS(file.path(results.dir, "cross_pairscan.RDS"))
}
```

## Motifs

First we look at motifs in this network. There is no enrichment of any 
type of motif.

```{r motif_enrichment, fig.width = 7, fig.height = 5}
motif.file <- here(results.dir, "Motif.Results.RDS")

if(!file.exists(motif.file)){
  motif.obj <- find.motifs(final_obj)
  motif.count <- count.motifs(motif.obj)
  motif.enrich <- motif.enrichment(final_obj)
  saveRDS(motif.enrich, motif.file)
}else{
  motif.enrich <- readRDS(motif.file)
}

#jpeg("~/Desktop/motifs.jpg", width = 7, height = 5, units = "in", res = 300)
fig.cols <- c(1:3, 6:8)
par(mar = c(8, 4, 4, 2))
main.shift = 0.3; col.text.shift = 0.27; row.text.shift = 0.3; col.text.rotation = 45
imageWithText(motif.enrich$overlayed[,fig.cols], 
    col.names = colnames(motif.enrich$motif.counts)[fig.cols], 
			row.names = rownames(motif.enrich$motif.counts), 
      main = "Overlay of Enriched (brown) and Depleted (blue) Motifs", 
			class.mat = motif.enrich$overlayed.class.mat[,fig.cols], 
      grad.dir = "low", color.fun = "exponential", exp.steepness = 10, 
			global.color.scale = FALSE, global.min = 0, global.max = 1, cex = 1.2, 
			col.text.adj = 1, main.shift = main.shift, col.text.shift = col.text.shift, 
			row.text.shift = row.text.shift, col.text.rotation = col.text.rotation,
      col.scale = c("brown", "blue"), light.dark = "l")	 	
#dev.off()
```

## Allele Interaction Patterns

The following code focuses on strain effects both 
for main effects and interactions.

```{r strain_main_effects}
all.inf <- write_variant_influences(final_obj, p_or_q = 0.05, write_file = FALSE)
main.locale <- which(all.inf[,"Target"] %in% colnames(final_obj$pheno))
just.main <- all.inf[main.locale,]
main.marker <- just.main[,"Source"]
split.marker <- strsplit(main.marker, "_")
just.marker <- sapply(split.marker, function(x) x[1])
just.allele <- sapply(split.marker, function(x) x[2])
```

The following bar plot shows the number of main effects derived
from each strain.

```{r plot_num}
barplot(table(just.allele), ylab = "Number of Main Effects")
```

The following boxplot shows the sign and magnitude of these
main effects for each trait.

```{r box_main, fig.width = 8, fig.height = 5}
u_alleles = sort(unique(just.allele))
u_pheno <- colnames(final_obj$pheno)
pheno_effects <- vector(mode = "list", length = length(u_pheno))
names(pheno_effects) <- u_pheno
for(i in 1:length(u_pheno)){
  allele_effects <- vector(mode = "list", length = length(u_alleles))
  names(allele_effects) <- u_alleles
  pheno.locale <- which(just.main[,"Target"] == u_pheno[i])
  for(j in 1:length(u_alleles)){
    allele.locale <- which(just.allele == u_alleles[j])
    overlap.locale <- intersect(allele.locale, pheno.locale)
    allele_effects[[j]] <- as.numeric(just.main[overlap.locale,"Effect"])
  }
  pheno_effects[[i]] <- allele_effects
}
plot.grouped.boxes(pheno_effects, legend.x = 17, legend.y = -0.1)
abline(h = 0)
```

The following heat map shows how many interactions were 
identified between each of the allele pairs.

```{r allele_int}
just.int <- all.inf[which(is.na(all.inf[,"conditioning_marker"])),]
source.marker <- just.int[,"Source"]
target.marker <- just.int[,"Target"]
split.source <- strsplit(source.marker, "_")
source.allele <- sapply(split.source, function(x) x[2])
split.target <- strsplit(target.marker, "_")
target.allele <- sapply(split.target, function(x) x[2])

allele.int <- cbind(source.allele, target.allele)
int.count <- as.matrix(table(data.frame(allele.int)))
imageWithText(int.count, col.scale = "blue", row.text.shift = 0.1,
col.text.shift = 0.12, col.text.rotation = 0, cex = 1.5)

#int.net <- graph_from_edgelist(allele.int)
#plot(int.net)
```

```{r combo_effects}
source.effect <- vector(mode = "list", length = length(u_alleles))
names(source.effect) <- u_alleles
mean.effect <- matrix(NA, nrow = 8, ncol = 8)
rownames(mean.effect) <- colnames(mean.effect) <- 1:8
for(sallele in u_alleles){
  target.effect <- vector(mode = "list", length = length(u_alleles))
  names(target.effect) <- u_alleles
  source.locale <- which(source.allele == sallele)
  for(tallele in u_alleles){
    target.locale <- which(target.allele == tallele)
    overlap.locale <- intersect(source.locale, target.locale)
    target.effect[[tallele]] <- abs(as.numeric(just.int[overlap.locale,"Effect"]))
    mean.effect[sallele,tallele] <- median(as.numeric(just.int[overlap.locale,"Effect"]))
  }
  source.effect[[sallele]] <- target.effect
}
```

The following boxplot shows the distribution of effects 
of interactions between each pair of alleles. The effects
are grouped first by source allele then by target allele.

```{r plot_boxes}
plot.grouped.boxes(source.effect)
abline(h = 0)
```

The following heat map shows the median effect across 
all interacting pairs of alleles. Interactions within
strain are along the diagonal.

```{r plot_mat}
imageWithText(round(mean.effect, 2), col.scale = c("blue", "brown"), 
split.at.vals = TRUE, row.text.shift = 0.1, grad.dir = "ends",
col.text.shift = 0.12, col.text.rotation = 0, cex = 1.5)
```

The following boxes show the magnitude of effects of interactions
within strain and between strain.

```{r agg_plot}
boxplot(list(abs(diag(mean.effect)), c(abs(mean.effect[upper.tri(mean.effect)]), 
  abs(mean.effect[lower.tri(mean.effect)]))), names = c("Within", "Between"))
```

The only pattern I can see is that alleles with more main effects
also have more interactions.